Supplementary Material

### 2021-10-22 analysis/paper/paper.Rmd
remove redundant code from paper and R scripts

```{r solution-count-out}
# sol_comb_row <- sol_comb_out %>%
#   group_by(treatment) %>%
#   filter(treatment == "mix") %>%
#   summarise(starch = "both",
#             across(where(is.numeric), sum, na.rm = T))

sol_tbl <- sol_comb_out %>%
  #ungroup() %>%
  #add_row(sol_comb_row, .after = 2) %>%
  mutate(across(c(s, m, l, total), 
                function(x) 
                  paste0(round(x), " (", scales::percent(x / total, 0.1), ")")))

# count_sol_out <- count_sol %>%
#   mutate(across(c(s, m, l, total), 
#                 function(x) 
#                   paste0(round(x), " (", scales::percent(x / total, 0.1), ")")))
# 
# ### adding the 'both' row to the data frame ###
# count_sol_comb_out <- count_sol_comb %>%
#   mutate(across(c(s, m, l, total), 
#                 function(x) 
#                   paste0(round(x), " (", scales::percent(x / total, 0.1), ")")))
# 
# count_sol_out <- count_sol %>%
#   mutate(across(c(s, m, l, total), 
#                 function(x) 
#                   paste0(round(x), " (", scales::percent(x / total, 0.1), ")"))) %>%
#   ungroup() %>%
#   add_row(tibble_row(count_sol_comb_out), .before = 3)

#count_sol_out[1,3] <- NA
sol_tbl[1,3] <- NA
```

@@

# count_sample <- count_sample %>%
#   ungroup() %>%
#   add_row(samp_sd_row, .after = 2)

# convert columns to contain mean count and percentage of total count
# count_samp_per <- count_sample %>%
#   mutate(across(c(s, m, l, total), 
#                 function(x) 
#                   paste0(round(x), " (", scales::percent(signif(x / total, 3)), ")")))

# count_samp_out <- tibble(count_samp_per, count_sample_sd, .name_repair = "unique") %>%
#   select(1,2,3,7,4,8,5,9,6,10) #%>%

samp_counts_tbl[1,3] <- NA
#count_samp_out[1,3] <- NA # replace with NA
  #add_row(tibble_row(count_sample_comb_out), .before = 3)
```

It was not possible to differentiate between potato and 
wheat starches smaller than ca. 10 $\mu$m. These were counted as wheat,
as we assumed that the majority of the small granules were wheat. We make this 
assumption based on the counts of small starches in the wheat-only and potato-only
solutions. Of the total amount of small starches in these two solutions, 
`r paste0(signif((sol_comb_out$s[4] / sum(sol_comb_out$s[3], sol_comb_out$s[4])) * 100, 3), "%")`
are from wheat.
<!-- count of small granules from wheat solution divided by the sum of small granules 
from the potato and wheat solutions -->

@@

```{r sample-perc}
# convert proportions to percent labels
# perc_comb <- prop_comb %>% 
#   mutate(across(c(s, m, l, total), scales::percent, 0.001))

# perc <- prop %>%
#   mutate(across(c(s, m, l, total), scales::percent, 0.001)) %>% 
#   add_row(tibble_row(perc_comb), .before = 3)

prop_wide <- prop %>%
  ungroup() %>%
  pivot_wider(id_cols = c(treatment, starch), 
              names_from = size, 
              values_from = prop) %>%
  add_row(prop_comb_row, .after = 2)

perc_tbl <- prop %>%
  ungroup() %>%
  select(!c(sample, solution, prop)) %>%
  pivot_wider(names_from = size, values_from = perc) %>%
  add_row(perc_comb_row, .after = 2)
# samp_comb_row <- corr_comb_long %>%
#   pivot_wider(names_from = size, values_from = count) %>%
#   group_by(treatment) %>%
#   filter(treatment == "mix") %>%
#   summarise(starch = "both",
#             across(where(is.numeric), sum, na.rm = T)) %>%
#   mutate(sample = "sample")

# perc_comb_row <- sol_comb_row %>%
#   mutate(sample = "solution") %>%
#   add_row(samp_comb_row) %>%
#   pivot_longer(cols = c(s, m, l, total), names_to = "size")
#   mutate()

#perc <- prop %>%
#  pivot_wider(id_cols = c(treatment, starch), names_from = size, values_from = perc)
```

>>>>

```{r solution-count-out}
sol_tbl <- sol_comb_out %>%
  mutate(across(c(s, m, l, total), 
                function(x) 
                  paste0(round(x), " (", scales::percent(x / total, 0.1), ")")))

sol_tbl[1,3] <- NA
```

@@

samp_counts_tbl[1,3] <- NA
```

It was not possible to differentiate between potato and 
wheat starches smaller than ca. 10 $\mu$m. These were counted as wheat,
as we assumed that the majority of the small granules were wheat. We make this 
assumption based on the counts of small starches in the wheat-only and potato-only
solutions. Of the total amount of small starches in these two solutions, 
`r paste0(signif((sol_comb_out$s[5] / sum(sol_comb_out$s[4], sol_comb_out$s[5])) * 100, 3), "%")`
are from wheat.
<!-- count of small granules from wheat solution divided by the sum of small granules 
from the potato and wheat solutions -->

@@

```{r sample-perc}
prop_wide <- prop %>%
  ungroup() %>%
  pivot_wider(id_cols = c(treatment, starch), 
              names_from = size, 
              values_from = prop) %>%
  add_row(prop_comb_row, .after = 2)

perc_tbl <- prop %>%
  ungroup() %>%
  select(!c(sample, solution, prop)) %>%
  pivot_wider(names_from = size, values_from = perc) %>%
  add_row(perc_comb_row, .after = 2)
```

### 2021-10-22 analysis/supplementary-materials/supp-mat.Rmd
remove redundant code from paper and R scripts

---
title: Supplementary Materials
output: 
  html_document:
    keep_md: true
    df_print: paged
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(here)
library(tidyverse)
raw_data_samples <- read_csv(here("analysis/data/raw_data/starch_counts.csv"))
raw_data_solutions <- read_csv(here("analysis/data/raw_data/solution_counts.csv"))
```

@@

```{r plate-lid-2, fig.cap="The 24 deepwell plate with the lid on (sort of...)"}
knitr::include_graphics(here("analysis/figures/plate_lid_on.jpg"))
```

@@

**solution_counts**

@@

```{r raw-data-sol, echo=FALSE}
raw_data_solutions
```

@@

```{r raw-data-samp, echo=FALSE}
raw_data_samples
```

@@

Tables containing the amylase activity results for both plates and photometric
readings.

```{r amylase-activity-results, message=FALSE}
# table of results reported in units amylase per mL enzyme (but let's be honest,
  # ...it doesn't really matter what the unit is. No activity is no activity)
source(here("analysis/scripts/amylase-assay.R"))
cols <- c("1", "2", "3")  
rows <- c("S1", "S2", "S3", "B1", "B2", "B3", "B4", "B5", "BT1", "BT2", "BT3")
plt1_ph1_result <- rbind(sal1_ph1, bmm1_ph1)
rownames(plt1_ph1_result) <- rows
plt1_ph2_result <- rbind(sal1_ph2, bmm1_ph2)
rownames(plt1_ph2_result) <- rows
plt2_ph1_result <- rbind(sal2_ph1, bmm2_ph1)
rownames(plt2_ph1_result) <- rows
plt2_ph2_result <- rbind(sal2_ph2, bmm2_ph2)
rownames(plt2_ph2_result) <- rows
```

@@

```{r}
raw_data_samples %>%
  filter(treatment == "control") %>%
  select(!c(vol, portion_slide, s, m, l))
```

@@

Solution counts were extrapolated to the quantity in a 1 ml solution with
the following formula:

$$ \text{corrected count} = \text{raw count} \times \frac{\text{total} \space \text{transects}}{\text{counted} \space \text{transects}} \times 100 \mu l $$

Sample counts were extrapolated using the following formula:

@@

>>>>

---
title: Supplementary Material
output: 
  html_document:
    keep_md: true
    df_print: paged
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(patchwork)
source(here("analysis/scripts/analysis.R"))
#raw_data_samples <- read_csv(here("analysis/data/raw_data/starch_counts.csv"))
#raw_data_solutions <- read_csv(here("analysis/data/raw_data/solution_counts.csv"))
```

@@

```{r plate-lid-2, fig.cap="The 24 deepwell plate with the lid (almost) on."}
knitr::include_graphics(here("analysis/figures/plate_lid_on.jpg"))
```

@@

**solution_counts.csv**

@@

```{r raw-data-sol, echo=FALSE}
sol_counts
```

@@

```{r raw-data-samp, echo=FALSE}
raw_counts
```

@@

Tables containing the amylase activity results for both plates and both photometric
readings conducted on each plate. Samples (rows) were analysed in triplicates (columns).

```{r amylase-activity-results, message=FALSE}
# table of results reported in units amylase per mL enzyme (but let's be honest,
  # ...it doesn't really matter what the unit is. No activity is no activity)
source(here("analysis/scripts/amylase-assay.R"))
cols <- c("1", "2", "3")  # sample triplicates
rows <- c("S1", "S2", "S3", "B1", "B2", "B3", "B4", "B5", "BT1", "BT2", "BT3")
plt1_ph1_result <- rbind(sal1_ph1, bmm1_ph1)
rownames(plt1_ph1_result) <- rows
plt1_ph2_result <- rbind(sal1_ph2, bmm1_ph2)
rownames(plt1_ph2_result) <- rows
plt2_ph1_result <- rbind(sal2_ph1, bmm2_ph1)
rownames(plt2_ph1_result) <- rows
plt2_ph2_result <- rbind(sal2_ph2, bmm2_ph2)
rownames(plt2_ph2_result) <- rows
```

@@

```{r}
raw_counts %>%
  filter(treatment == "control") %>%
  select(!c(vol, portion_slide, s, m, l))
```

@@

A 1 mL aliquot of each of the original treatment solutions was taken, from which
10 $\mu$L was taken and mounted on a microscope slide and mixed with 10 $\mu$L 
20% (v/v) glycerol. Solution counts were extrapolated from a slide (10 $\mu$L) 
to the quantity in a 1 ml solution,
and then multipled by 16 days to achieve the total number of granules that were
exposed to the samples:

$$ \text{corrected count} = \text{raw count} \times \frac{\text{total slides}}{\text{counted slides}} \times 100 \mu L \times 16 \text{ days} $$

Samples were submerged in 50--100 $\mu$L EDTA, from which 20 $\mu$L was mounted 
on a microscope slide ($V_{slide}$) and counted. Sample counts were extrapolated to the full
volume of EDTA ($V_{sample}$) in which the sample was submerged (i.e. 50--100 $\mu$L).

@@

## Some additional plots

Bar plot for the total count of granules exposed to the samples over the
duration of the experiment,

```{r solution-bar-pl}
sol_comb_long %>%
  filter(size == "total") %>%
  group_by(treatment, starch) %>%
  ggplot(aes(x = treatment, y = count, fill = treatment, col = starch)) +
  geom_col(size = 1.5) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "grey"),
        panel.grid.major.x = element_blank(),
        axis.title.x = element_blank()) +
  scale_fill_viridis_d() +
  scale_color_viridis_d(begin = 0.5)
```

and box plot with superimposed points (with added jitter) for the extrapolated
mean counts of granules extracted from the samples.

```{r sample-box-plot}
corr_comb %>%
  filter(treatment != "control") %>%
  ggplot(aes(x = treatment, y = total, 
             shape = treatment)) +
    geom_boxplot(aes(fill = treatment), alpha = 0.5) +
    geom_jitter(aes(col = treatment), width = 0.3, size = 2) +
    scale_color_viridis_d() +
    theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_line(colour = "grey"),
        panel.grid.major.x = element_blank(),
        axis.title.x = element_blank()) + # remove y-axis title
    scale_fill_viridis_d()
```

Extracted-granule counts separated by size:

```{r}
corr_counts_long %>%
  filter(size != "total",
         treatment != "control") %>%
  group_by(treatment, starch, size) %>%
  summarise(sd = sd(count, na.rm = T),
            count = mean(count, na.rm = T)) %>%
  #mutate(percent = count / sum(count, na.rm = T) * 100) %>% 
  ggplot(aes(x = starch, y = count, fill = size)) +
    geom_col(position = "dodge") +
    geom_errorbar(aes(ymin = count, ymax = count + sd), width = 0.2, position = position_dodge(0.9)) +
    facet_wrap(~ treatment, scales = "free") +
    scale_fill_viridis_d() +
    theme_bw()
```

Size distribution within the solutions (top) and samples (bottom):

```{r message=FALSE, warning=FALSE}
sol_size_pl <- sol_corr %>%
  group_by(solution, starch) %>%
  summarise(across(c(s, m, l, total), mean, na.rm = T)) %>%
  pivot_longer(cols = c(s,m,l, total), values_to = "count", names_to = "size") %>%
  filter(size != "total") %>%
  group_by(solution, starch) %>%
  mutate(percent = count / sum(count, na.rm = T) * 100) %>% 
  ggplot(aes(x = starch, y = percent, fill = size)) +
    geom_col(position = "dodge") +
    facet_wrap(~ solution, scales = "free_x") +
    scale_fill_viridis_d() +
    theme_bw() +
    labs(x = "")

samp_size_pl <- corr_counts_long %>%
  filter(size != "total",
         treatment != "control") %>%
  group_by(treatment, starch, size) %>%
  summarise(count = mean(count, na.rm = T)) %>%
  mutate(percent = count / sum(count, na.rm = T) * 100) %>% 
  ggplot(aes(x = starch, y = percent, fill = size)) +
    geom_col(position = "dodge") +
    facet_wrap(~ treatment, scales = "free_x") +
    scale_fill_viridis_d() +
    theme_bw()

sol_size_pl / samp_size_pl + plot_layout(guides = "collect")
```

Size distribution within the samples:

```{r message=FALSE, warning=FALSE}
samp_size_pl <- corr_counts_long %>%
  filter(size != "total",
         treatment != "control") %>%
  group_by(treatment, starch, size) %>%
  summarise(count = mean(count, na.rm = T)) %>%
  mutate(percent = count / sum(count, na.rm = T) * 100) %>% 
  ggplot(aes(x = starch, y = percent, fill = size)) +
    geom_col(position = "dodge") +
    facet_wrap(~ treatment, scales = "free_x") +
    scale_fill_viridis_d() +
    theme_bw()
```
